---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# integrateIt

<!-- badges: start -->
<!-- badges: end -->

The goal of integrateIt is to approximate numerical integrations using the Trapezoid Rule and Simpson's Rule.

## Installation

You can install the development version of integrateIt from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("johnsontr/AppliedStatisticalProgramming2022")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(integrateIt)
f <- function(x) {sin(x)}
simpson_fun(a = 0, b = 10, d = 5, f)
trapezoid_fun(a = 0, b = 10, d = 5, f)
```

The functions will output a numerical estimation for the definite integral approximations.

The package works by taking in arguments:
a = the start value for an integral to be estimated
b = the end value for an integral to be estimated
d = the number of divisions to be estimated
f = the function being integrated

It passes these arguments through the functions trapezoid_fun or simpson_fun, depending on whether the user wants to use the Trapezoid rule or the Simpson rule for approximation, respectively.
These functions output objects of their respective class, Trapezoid or Simpson. The outputs include x (the vector of values that were to be evaluated), y (the vector evaluations), and the approximated result.
The getResult, getX, and getY methods access the results, x, and y, individually from each class.
integrateIt uses the getResult, getX, and getY methods to output the x, y, and results of the rule approximations.
print uses getResult to print the name of the rule being utilized and the approximated result.
